---
title: "Column Operations"
date: 05-27-2023
categories: [R] # self-defined categories
draft: false # setting this to \`true\` will prevent your post from appearing on your listing page until you're ready!
---

#### Renaming Columns

The two functions I use to rename columns are dplyr::rename() and dplyr::rename_with()

rename() is good for when you need to manually code in the new name

```{r}
#| eval: false 
sampledf <- sampledf |> 
  rename(
    # NEW = OLD
    country     = country_name,
    supply_plan = supply_plan_description
  )
```

rename_with() is for when you need to use a function to rename columns. For example if you want to pick all columns that end with "\_supply_plan", remove that suffix and add the prefix "value\_" this is what you would do:

```{r}
#| eval: false 

sampledf <- sampledf |> 
  rename_with(
    ~ str_c("value_",
             str_remove(.x,"_supply_plan")
    ),
    .cols = ends_with("_supply_plan")
  )
```

Note the use of the [tidyselect](https://tidyselect.r-lib.org/reference/language.html) ends_with instead of base R endswith. No difference in this case between the two but the tidyselect function is more versatile as it includes an ignore.case argument.

Older code often has paste or paste0 to concatenate strings. Here I use stringr::str_c instead.

#### **Relocating columns**

can be done with dplyr::relocate() with [tidy-select](https://dplyr.tidyverse.org/reference/dplyr_tidy_select.html).

#### **Removing columns**

is easy with dplyr::select(! column_name)

#### Applying a function to multiple columns

is done using dplyr::across()

```{r}
#|eval:false

sampledf <- sampledf |> 
  mutate(across(.cols = im_number:im_funding, 
                .fns = \(x) str_remove_all(x, "@|:@")),
         
         across(.cols = where(is.POSIXct), 
                .fns = \(x) as_date(x))
  )
            
```

Note the use of \\(x) instead of the older \~. to start an unnamed function
